---
title: Canvas Options
---

```{r}
library(gglite)
```

The `canvas()` function controls chart dimensions, layout spacing, and the
rendering backend. Pipe it after `g2()` to customize a chart's canvas before
rendering.

## Dimensions

Use `width` and `height` to set the chart size in pixels. By default (`width =
NULL`), the chart auto-fits to its container width and uses a height of 480 px.

```{r}
g2(mtcars, hp ~ mpg) |> canvas(width = 500, height = 300)
```

## Layout Spacing

The `padding`, `margin`, and `inset` arguments control spacing in pixels. Each
accepts a scalar (all sides) or a length-4 vector `c(top, right, bottom, left)`
where `NA` skips a side.

```{r}
# Add padding around the plot area (useful for parallel coords, etc.)
g2(iris, position = names(iris)[-5], color = ~ Species) |>
  canvas(padding = 30) |>
  coord_parallel() |>
  legend_color(position = 'bottom')
```

## Renderer

G2 supports three rendering backends, selected via the `renderer` argument
(case-insensitive):

| Renderer | CDN scripts loaded | Approx. size |
|---|---|---|
| `"Canvas"` (default) | `g2.min.js` | ~1.1 MB |
| `"SVG"` | `@antv/g` + `@antv/g-svg` + `g2.lite.min.js` | ~1.5 MB |
| `"WebGL"` | `@antv/g` + `@antv/g-webgl` + `g2.lite.min.js` | ~1.9 MB |

The default Canvas renderer is the lightest option. SVG output can be inspected
in the browser's DevTools (useful for debugging). WebGL is GPU-accelerated and
suited for very large datasets.

```{r}
# SVG renderer — standalone preview, no global option needed
g2(mtcars, hp ~ mpg) |> canvas(renderer = 'svg')
```

### Caveat: using SVG or WebGL in a multi-chart document

`g2.min.js` (Canvas) and `g2.lite.min.js` (SVG/WebGL) **cannot coexist on the
same HTML page**. In a document that contains multiple charts — such as an R
Markdown, Quarto, or Jupyter notebook — the CDN scripts are injected once for
the whole document. If any chart uses SVG or WebGL, all charts must use the
lite bundle.

To switch the entire document to the lite bundle, declare a global renderer
option **at the top of the document** (before any charts are rendered):

```r
options(gglite.renderer = 'svg')   # or 'webgl'
```

When this option is set:

- All charts in the document use `g2.lite.min.js` automatically.
- Individual charts can still override the renderer via `canvas(renderer = ...)`.
  For example, when `options(gglite.renderer = 'svg')` is active, a specific
  chart can use `g2(...) |> canvas(renderer = 'webgl')`.

For **standalone plots** previewed in the browser (`print()` / RStudio Viewer),
each plot is its own HTML page, so no global option is needed.

```{r echo=FALSE}
# Set global renderer to svg for the rest of this document
options(gglite.renderer = 'svg')
```

```{r}
# With options(gglite.renderer = 'svg') set, all charts use g2.lite
g2(mtcars, hp ~ mpg)
```

```{r}
# Override to WebGL for a specific chart
g2(mtcars, hp ~ mpg) |> canvas(renderer = 'webgl')
```

```{r echo=FALSE}
options(gglite.renderer = NULL)  # restore default
```

## Extra Options

The `...` argument in `canvas()` accepts any additional top-level
`chart.options()` keys (e.g., `clip`, `depth`) that are not covered by the
dedicated parameters.

```{r}
g2(mtcars, hp ~ mpg) |> canvas(clip = TRUE)
```
