---
title: Data
---

```{r}
library(gglite)
```

gglite automatically trims dataset columns to only those referenced by the
chart before serializing to JSON. This keeps the HTML output compact when
working with wide data frames that have many unused columns.

## Automatic Column Trimming

The `iris` dataset has five columns: `Sepal.Length`, `Sepal.Width`,
`Petal.Length`, `Petal.Width`, and `Species`. When only two columns are
mapped, only those two columns end up in the generated HTML.

```{r}
# Only Sepal.Length and Sepal.Width are included in the output
g2(iris, x = 'Sepal.Length', y = 'Sepal.Width') |> mark_point()
```

Additional aesthetic channels count as used columns too:

```{r}
# Sepal.Length, Sepal.Width, and Species are included; Petal.* are dropped
g2(iris, x = 'Sepal.Length', y = 'Sepal.Width', color = 'Species') |>
  mark_point()
```

Trimming also applies to data supplied at the mark level:

```{r}
df = data.frame(x = 1:5, y = c(2, 4, 3, 5, 1), z = letters[1:5])
# z is not encoded, so it is excluded from the mark's data
g2() |> mark_point(data = df, encode = list(x = 'x', y = 'y'))
```

## Opting Out with `I()`

Some chart configurations reference columns inside inline JavaScript
functions that gglite cannot detect statically. A common case is a custom
`style` callback that reads a column from the data row `d` directly:

```{r}
# Species is used in the JS fill function but is not listed in the encode.
# Without I(), Species would be trimmed and the colors would all fall back
# to the else branch.
g2(I(iris), x = 'Sepal.Length', y = 'Sepal.Width') |>
  mark_point(style = list(
    fill = js('(d) => d.Species === "setosa" ? "steelblue" : "tomato"')
  ))
```

Wrapping the data in `I()` tells gglite to preserve all columns and skip
trimming. The `AsIs` class is stripped before JSON serialization, so the
chart works exactly as if the data were passed directly — but with all
columns available to JavaScript.

The same applies to mark-level data:

```{r}
df = data.frame(x = 1:5, y = c(2, 4, 3, 5, 1), label = c('a', 'b', 'c', 'd', 'e'))
# label is referenced in a JS tooltip callback, not in encode
g2() |> mark_point(
  data = I(df),
  encode = list(x = 'x', y = 'y'),
  tooltip = list(items = list(js('(d) => ({ name: "label", value: d.label })')))
)
```
