---
title: Titles and Tooltips
---

```{r}
library(gglite)
p = g2(mtcars, hp ~ mpg)
```

## Chart Title

### Title and subtitle via `g2()` arguments

```{r}
g2(mtcars, hp ~ mpg,
  title = 'Motor Trend Cars', subtitle = 'mpg vs horsepower')
```

### Simple title

```{r}
p |> title_('Motor Trend Cars')
```

### Title with subtitle

```{r}
p |> title_('Motor Trend Cars', subtitle = 'mpg vs horsepower')
```

### Title alignment

```{r}
p |> title_('Right-aligned Title', align = 'right')
```

### Title on a bar chart

```{r}
df = data.frame(x = c('A', 'B', 'C', 'D'), y = c(3, 7, 2, 5))
g2(df, y ~ x) |>
  title_('Simple Bar Chart', subtitle = 'Categories A-D')
```

## Tooltip Configuration

### Enable crosshairs

Crosshairs are lines that track the cursor position. Pass `crosshairs = TRUE`
to `tooltip_()`. Crosshairs work with line and area charts (which use series
tooltip by default).

```{r}
air = aggregate(Temp ~ Month, data = airquality, FUN = mean)
air$Month = month.abb[air$Month]
g2(air, Temp ~ Month) |>
  mark_line() |>
  mark_point() |>
  tooltip_(crosshairs = TRUE)
```

### Crosshairs with custom style

```{r}
g2(air, Temp ~ Month) |>
  mark_line() |>
  tooltip_(crosshairs = TRUE, crosshairsStroke = '#aaa', crosshairsLineWidth = 1)
```

### Shared tooltip (useful for multi-series)

A shared tooltip shows values from all series at the current x position.

```{r}
df = data.frame(
  x = rep(1:5, 2), y = c(3, 1, 4, 1, 5, 2, 7, 1, 8, 3),
  group = rep(c('A', 'B'), each = 5)
)
g2(df, y ~ x, color = ~ group) |>
  mark_line() |>
  tooltip_(shared = TRUE)
```

### Format tooltip values

Pass a `tooltip` list to the mark function to format values using a
[d3-format](https://d3js.org/d3-format) string. This keeps the tooltip data
configuration with the mark that owns it.

```{r}
g2(airquality, Temp ~ Wind) |>
  mark_point(tooltip = list(channel = 'y', valueFormatter = '.1f'))
```

### Format tooltip values in a normalized stacked bar chart

After `transform_('normalizeY')`, y values are on a 0–1 scale. Use
`valueFormatter = '.0%'` to display them as percentages.

```{r}
wb = aggregate(breaks ~ wool + tension, data = warpbreaks, FUN = mean)
g2(wb, breaks ~ tension, color = ~ wool) |>
  mark_interval(tooltip = list(channel = 'y', valueFormatter = '.0%')) |>
  transform_('stackY') |>
  transform_('normalizeY') |>
  axis_y(labelFormatter = '.0%') |>
  legend_color(position = 'right') |>
  title_('Warp Breaks by Tension and Wool Type')
```

### Disable tooltip

```{r}
p |> tooltip_(FALSE)
```

