---
title: Tooltips
---

```{r}
library(gglite)
```

Tooltips in gglite have two distinct configuration points that match G2's
internal structure:

- **Chart-level** (`tooltip_()` → `interaction.tooltip`): controls tooltip
  *behavior* — crosshairs, shared mode, marker style, position, and rendering.
- **Mark-level** (`mark_*(tooltip = list(...))` → mark spec `tooltip`): controls
  what *data* is shown for a specific mark — which channels to display, value
  formatting, custom item names.

## Chart-Level Tooltip (`tooltip_()`)

All arguments to `tooltip_()` are passed to G2's `interaction.tooltip`. These
options apply across all marks in the chart.

### Disable Tooltip

Pass `FALSE` to turn off the tooltip entirely.

```{r}
g2(mtcars, hp ~ mpg) |> tooltip_(FALSE)
```

### Crosshairs

Crosshairs are reference lines that track the cursor. They work with marks that
use the *series* tooltip path (line, area, etc.).

```{r}
air_monthly = aggregate(Temp ~ Month, data = airquality, FUN = mean)
air_monthly$Month = month.abb[air_monthly$Month]
g2(air_monthly, Temp ~ Month) |>
  mark_line() |>
  mark_point() |>
  tooltip_(crosshairs = TRUE)
```

### Crosshairs on One Axis Only

`crosshairsX = TRUE` shows only the vertical line; `crosshairsY = TRUE` shows
only the horizontal line.

```{r}
g2(air_monthly, Temp ~ Month) |>
  mark_line() |>
  tooltip_(crosshairsX = TRUE)
```

```{r}
g2(air_monthly, Temp ~ Month) |>
  mark_line() |>
  tooltip_(crosshairsY = TRUE)
```

### Crosshairs Style

Use `crosshairsStroke`, `crosshairsLineWidth`, and `crosshairsLineDash` to
customize the appearance of crosshairs.

```{r}
g2(air_monthly, Temp ~ Month) |>
  mark_line() |>
  tooltip_(
    crosshairs = TRUE,
    crosshairsStroke = '#999',
    crosshairsLineWidth = 1,
    crosshairsLineDash = c(4, 4)
  )
```

### Shared Tooltip

`shared = TRUE` shows a single tooltip listing all mark values at the same x
position. Useful for comparing multiple series.

```{r}
df = data.frame(
  x = rep(1:5, 2), y = c(3, 1, 4, 1, 5, 2, 7, 1, 8, 3),
  group = rep(c('A', 'B'), each = 5)
)
g2(df, y ~ x, color = ~ group) |>
  mark_line() |>
  tooltip_(shared = TRUE)
```

### Marker Style

The hover marker is the dot that appears on the data point. Use `marker = FALSE`
to hide it, or customize with `markerType`, `markerFill`, `markerStroke`, and
`markerSize`.

```{r}
# Hide the hover marker
g2(air_monthly, Temp ~ Month) |>
  mark_line() |>
  tooltip_(marker = FALSE)
```

```{r}
# Hollow marker
g2(air_monthly, Temp ~ Month) |>
  mark_line() |>
  tooltip_(markerType = 'hollow', markerSize = 8)
```

### Hide Tooltip Body

`body = FALSE` shows only the hover marker with no tooltip box.

```{r}
g2(air_monthly, Temp ~ Month) |>
  mark_line() |>
  tooltip_(body = FALSE)
```

### Tooltip Position

Control where the tooltip appears relative to the cursor. Possible values:
`'top'`, `'bottom'`, `'left'`, `'right'`, `'top-left'`, `'top-right'`,
`'bottom-left'`, `'bottom-right'` (default: `'right-bottom'`).

```{r}
g2(mtcars, hp ~ mpg) |>
  tooltip_(position = 'top-left')
```

### Tooltip Offset

Adjust the tooltip's x/y offset from the cursor in pixels (default: `c(10, 10)`).

```{r}
g2(mtcars, hp ~ mpg) |>
  tooltip_(offset = c(20, 20))
```

### Group Name

`groupName = FALSE` hides the series/group name prefix in each tooltip item.

```{r}
g2(df, y ~ x, color = ~ group) |>
  mark_line() |>
  tooltip_(shared = TRUE, groupName = FALSE)
```

### Enterable Tooltip

`enterable = TRUE` lets the user move the cursor inside the tooltip box
(useful when the tooltip contains clickable links or controls).

```{r}
g2(mtcars, hp ~ mpg) |>
  tooltip_(enterable = TRUE)
```

## Mark-Level Tooltip (`mark_*(tooltip = list(...))`)

Pass a `tooltip` argument to any `mark_*()` call to control what data is shown
in the tooltip for that specific mark. Options come from G2's `TooltipItem`
type.

### Show a Single Channel

Use `channel` to restrict the tooltip to one data channel.

```{r}
# Show only the y value
g2(airquality, Temp ~ Wind) |>
  mark_point(tooltip = list(channel = 'y'))
```

### Custom Item Name

Use `name` to label the tooltip item.

```{r}
g2(airquality, Temp ~ Wind) |>
  mark_point(tooltip = list(channel = 'y', name = 'Temperature (°F)'))
```

### Value Formatter

Use `valueFormatter` with a [d3-format](https://d3js.org/d3-format) string to
format the displayed value.

```{r}
# 1 decimal place
g2(airquality, Temp ~ Wind) |>
  mark_point(tooltip = list(channel = 'y', name = 'Temp', valueFormatter = '.1f'))
```

```{r}
# Percentage formatter in a normalized stacked bar chart
wb = aggregate(breaks ~ wool + tension, data = warpbreaks, FUN = mean)
g2(wb, breaks ~ tension, color = ~ wool) |>
  mark_interval(tooltip = list(channel = 'y', valueFormatter = '.2%')) |>
  transform_('stackY') |>
  transform_('normalizeY') |>
  axis_y(labelFormatter = '.0%') |>
  legend_color(position = 'right') |>
  title_('Warp Breaks by Tension and Wool Type')
```

### Custom Item Color

```{r}
g2(airquality, Temp ~ Wind) |>
  mark_point(tooltip = list(channel = 'y', name = 'Temp', color = 'steelblue'))
```

### Custom Title

Use `title` with a `channel` to display a specific data field as the tooltip
title.

```{r}
air_daily = airquality[, c('Month', 'Day', 'Temp')]
air_daily$Date = as.character(
  as.Date(paste('1973', air_daily$Month, air_daily$Day, sep = '-'))
)
g2(air_daily, Temp ~ Date) |>
  mark_line() |>
  mark_point(
    tooltip = list(
      title = list(channel = 'x'),
      items = list(list(channel = 'y', name = 'Temp (°F)'))
    )
  )
```

### Disable Tooltip for One Mark

Set `tooltip = FALSE` on a specific mark to hide its tooltip while keeping it
for others.

```{r}
air_monthly2 = aggregate(cbind(Temp, Wind) ~ Month, data = airquality, FUN = mean)
air_monthly2$Month = month.abb[air_monthly2$Month]

g2(air_monthly2, ~ Month) |>
  mark_interval(encode = list(y = 'Temp'), tooltip = FALSE) |>
  scale_y(independent = TRUE) |>
  axis_y(title = 'Temp (°F)') |>
  mark_line(encode = list(y = 'Wind')) |>
  scale_y(independent = TRUE) |>
  axis_y(position = 'right', title = 'Wind (mph)')
```
